Detection of Oscillations in Aldebaran with Ground-Based Observations

### 2018-01-05 paper/aldebaran.tex
satellites

In this paper, we present a re-analysis of these original RV data in which we not only confirm this signal, but detect acoustic oscillations in Aldebaran for the first time. We validate this method and its result with new independent RV observations with the SONG~Telescope, and photometry from the \ktwo Mission. By measuring the frequency of maximum power, \numax, of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun. This measured stellar mass allows us to calculate that any moons of this giant planet (S-type planets if Aldebaran~b is a brown dwarf), although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that they may have once been habitable.

@@

With the determination of the stellar mass it is possible to infer the parameters Aldebaran had while it was on the main sequence. We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathcal{N}(1.16,0.07)$ and metallicities $[\text{Fe}/\text{H}] \sim \mathcal{N}(-0.15,0.2)$ \citep{decin2003} to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.  %Using the scaling relations of \citep{eker2015}, we find that Aldebaran would have had a luminosity $L_* = 2.0 \pm 0.37~\lsun$ on the main sequence.  
Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models via the \textsc{isochrones} package \citep{isochrones}, we compute the evolution of the luminosity $L$ of the star along the main sequence; $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr. 
We note that this implies that the planet at a semi-major axis of $1.50\, \pm\, 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 4.5~Gyr from $0.86\, \pm\, 0.27$ to $1.55 \,\pm\,1.0$ times that of the Earth today (Figure~\ref{insolation}). Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, Aldebaran~b and any of its moons may well have hosted temperate environments for some of their history, now long-since destroyed by their star's evolution away from the main sequence.

>>>>

In this paper, we present a re-analysis of these original RV data in which we not only confirm this signal, but detect acoustic oscillations in Aldebaran for the first time. We validate this method and its result with new independent RV observations with the SONG~Telescope, and photometry from the \ktwo Mission. By measuring the frequency of maximum power, \numax, of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun. This measured stellar mass allows us to calculate that any satellites of Aldebaran~b, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that they may have once been habitable.

@@

With the determination of the stellar mass it is possible to infer the parameters Aldebaran had while it was on the main sequence. We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathcal{N}(1.16,0.07)$ and metallicities $[\text{Fe}/\text{H}] \sim \mathcal{N}(-0.15,0.2)$ \citep{decin2003} to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.  %Using the scaling relations of \citep{eker2015}, we find that Aldebaran would have had a luminosity $L_* = 2.0 \pm 0.37~\lsun$ on the main sequence.  
Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models via the \textsc{isochrones} package \citep{isochrones}, we compute the evolution of the luminosity $L$ of the star along the main sequence; $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr. 
We note that this implies that the planet at a semi-major axis of $1.50\, \pm\, 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 4.5~Gyr from $0.86\, \pm\, 0.27$ to $1.55 \,\pm\,1.0$ times that of the Earth today (Figure~\ref{insolation}). Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, Aldebaran~b and any of its moons (or its S-type planets if it is a brown dwarf) may well have hosted temperate environments for some of their history, now long-since destroyed by their star's evolution away from the main sequence.
